
---
  output: github_document
---
  
<!-- README.md is generated from README.Rmd. Please edit that file -->
  
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/",
  out.width = "100%"
)
```

# lmForc Helper Functions

```{r example100, message=FALSE}
library(lmForc)
```

Assorted functions for working with lmForc Forecast objects. The end game is to
integrate some of these functions into the lmForc package and turn the rest of
these functions into a standalone R package.

## Subsetting Example Dataset

Examples of lmForc subsetting functions utilize the following stylized dataset. 
This example dataset contains two forecasts with one quarter ahead forecasts
of the same variable. Note that both forecasts have identical `future`,
`realized`, and `h_ahead` values, but that `origin` dates of the last two forecasts
differ. These example forecasts can be thought of as two different model forecasts 
of the same variable.

```{r example101, message=FALSE}
forc1_1h <- Forecast(
  origin = as.Date(c("2010-02-17", "2010-05-14", "2010-07-22", "2010-12-05", "2011-03-10")),
  future = as.Date(c("2010-06-30", "2010-09-30", "2010-12-31", "2011-03-31", "2011-06-30")),
  forecast = c(4.27, 3.36, 4.78, 5.45, 5.12),
  realized = c(4.96, 4.17, 4.26, 4.99, 5.38),
  h_ahead = 1
)

forc2_1h <- Forecast(
  origin = as.Date(c("2010-02-17", "2010-05-14", "2010-07-22", "2010-12-22", "2011-03-27")),
  future = as.Date(c("2010-06-30", "2010-09-30", "2010-12-31", "2011-03-31", "2011-06-30")),
  forecast = c(4.01, 3.89, 3.31, 4.33, 4.61),
  realized = c(4.96, 4.17, 4.26, 4.99, 5.38),
  h_ahead = 1
)
```

## Subsetting Functions Overview

```{r example102, message=FALSE}
source("lmForc_subset.R")
```

The Forecast class has a built in method for subsetting a single Forecast object.
This subsetting method takes numeric or logical values and follows subsetting
conventions that are present throughout the R language.

```{r example103, message=FALSE}
forc1_1h[2:4]

forc1_1h[origin(forc1_1h) >= as.Date("2010-12-31")]
```

However, one often ends up working with multiple Forecast objects whether that be 
different model forecasts for the same forecast horizon, one model forecast for 
varying forecast horizons, or both. The lmForc convention for working with multiple
Forecast objects is to put them into a list. The following subsetting functios provide
a way to subset multiple Forecast objects by various conditions.

## subset_forcs

The simplest way to subset multiple Forecast objects is via the `subset_forcs()` 
function. The function takes a list of Forecast objects and a numeric or logical
index. All forecasts in the list are subset by the whatever is passed to the
`index` argument.

```{r example104, message=FALSE}
forcs <- list(forc1_1h, forc2_1h)

subset_forcs(forcs, 1:4)
 
subset_forcs(forcs, origin(forc1_1h) >= as.Date("2010-12-31"))
```

## subset_bytime

One may wish to compare forecasts over a specific time horizon. The `subset_bytime()`
allows the user to subset multiple forecasts based on origin or future values. After 
choosing whether to subset by origin or future values and passing a single time object
or a vector of time objects to the `values` argument, all forecasts in a list of 
Forecast objects are subset by `values`.

```{r example105, message=FALSE}
forcs <- list(forc1_1h, forc2_1h)
 
subset_bytime(forcs, values = as.Date("2010-05-14"), slot = "origin")

subset_bytime(
  forcs, 
  values = as.Date(c("2010-09-30", "2010-12-31", "2011-03-31")), 
  slot = "future"
)
```
 
## subset_identical

When comparing multiple forecasts, it is imperative that forecast accuracy is
compared across an identical time period. This can become complicated if the origin
and future values of multiple forecasts do not always align. The `subset_identical()`
function finds all overlapping origin or future values in a list of Forecast objects
and subsets each of the forecasts to these overlapping values. This leaves the user
with a list of Forecasts that have either identical origin values or identical 
future values depending on what the user passes to the `slot` argument.

```{r example106, message=FALSE}
forcs <- list(forc1_1h, forc2_1h)

subset_identical(forcs, slot = "origin")
```

## Transformation Functions Overview

```{r example107, message=FALSE}
source("lmForc_transform.R")
```

When compiling multiple Forecast objects, there are two general formats that a 
list of forecasts can take.

- time format (all forecasts have a homogenous origin or future)
- h_ahead format (all forecasts have a homogenous h_ahead)


# Helper Functions Overview

- Subset forecasts to identical future values.
- Then show how you can plot.
- Subset forecasts to specific future values,
- Then create RMSE table.






